### 2018-03-08 cuckoo_plots.Rmd
Initial update to project

---
title: "Cuckoo movements"
author: "MM"
date: "7 February 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cuckoos, echo=FALSE, warning=FALSE, message=FALSE}
########################

rm(list=ls())
library(maps)
library(sp)
library(maptools)
library(rgdal)
library(ggplot2)
library(ggmap)

# using best of day data
dat2<-read.csv("~/BTO/cuckoo_tracking/data/processed_movebank_cuckoos_hybrid_filter_bestofday_clean_stopovers.csv", h=T)

# hack to put months in correct numeric order
dat2$month2<-ifelse(dat2$month>9, dat2$month, paste(0, dat2$month, sep=""))
dat2$yrmo<-as.numeric(paste(dat2$year.x, dat2$month2, sep="."))

m1 <- get_map(location = "Nigeria", zoom=4)

for(i in unique(dat2$ptt))
{

  g<-ggmap(m1)+
    geom_line(data=dat2[dat2$ptt==i,], aes(x=long, y=lat,
                                             colour=factor(yrmo)))+
    geom_jitter(data=dat2[dat2$ptt==i,], aes(x=long, y=lat,
                colour=factor(yrmo)), shape=1, width = 1, height= 1)+ggtitle(i)

  print(g)
}

```