### 2018-05-30 stopover_rainfall_vis.Rmd
Now working with rainfall and ndvi

---
title: "West Africa cuckoo stopover rainfall patterns"
author: "Mark Miller"
date: "17 April 2018"
output: html_document
---

```{r cuckoos, echo=FALSE, message=F, warning=F, fig.height=12, fig.width=12}

#### work around for different laptop/desktop directories ####
#### !!!! RUN THIS !!!! ####
 if(Sys.info()['nodename']=="D9L5812"){
   setwd("C:/cuckoo_tracking")}else{
    setwd("N:/cuckoo_tracking")}

int_out<-read.csv('data/spring_rainfall_NDVI_by_stopover_2018_dead.csv', h=T)

dat<-read.csv("data/stopover_bestofday_2018_1daymin_recalc_spring_mig_summary_dead.csv", h=T)

### visualisastion

library(ggplot2)
library(dplyr)

# order country factors by longitude

int_out$country<-factor(int_out$country, levels= c("Angola", "Democratic Republic of the Congo",
                                                   "Congo", "Gabon", "Central African Republic", 
                                                   "Cameroon", "Nigeria", "Togo", "Burkina Faso",
                                                   "Ghana", "Cote d'Ivoire", "Liberia", "Guinea", 
                                                   "Sierra Leone"))
int_out[is.na(int_out$value),]$value<-0
int_out$ID2<-paste(int_out$SO_startDOY, int_out$country)

for( i in 2012:2017)
  {
  int1<-int_out[int_out$year==i,]

  for(j in unique(int1$ptt))
      {
dtemp<-dat[dat$year==i & dat$ptt==j,]

t1<-int_out[int_out$ptt==j & int_out$year==i,]

t2<-t1 %>% group_by(ID2) %>% dplyr::mutate(cumrf=cumsum(value))

temp=ggplot(data=t2,aes(x=variable, y=value))+
geom_rect(aes(xmin=SO_startDOY, xmax=SO_endDOY, ymin=0, ymax=500), fill='grey')+
    geom_bar(data=t2, aes(x=variable, y=value2*500, group=1), stat='identity', fill='red', linetype=0, alpha=1/3)+
  geom_bar(stat='identity', colour='black')+
    geom_line(aes(y=cumrf), col='green', size=1)+
  facet_wrap(~country+SO_startDOY, scales='free_y')+theme_classic()+
  ylab('cumulative rainfall since 01 Jan (mm)')+ xlab('Day of year')+
          ggtitle(paste("Year:",i,"   PTT:", j, "   Arr date:", dtemp$arrive_breeding))

#"   Mig speed:", round(dtemp$m4c_resid, 2)

print(temp)

#temp=ggplot(data=t2,aes(x=variable, y=value))+
#geom_rect(aes(xmin=SO_startDOY, xmax=SO_endDOY, ymin=0, ymax=50), #fill='grey')+
#  geom_bar(stat='identity', colour='black')+
#  geom_smooth(span=0.5, se=FALSE, linetype='dashed', col='red')+
#  facet_wrap(~country+SO_startDOY, scales='free_y')+theme_classic()+
#  ylab('daily rainfall (mm)')+ xlab('Day of year')+
#          ggtitle(paste("Year:",i,"   PTT:", j, "   Arr date:", #dtemp$arrive_breeding, "   Mig speed:", round(dtemp$m4c_resid, 2)))

  }
}

```