### 2018-08-30 stopover_rainfall_ndvi_detailcoords_vis_DEPWAbirds.Rmd
Minor change to specify just birds that we are using in DEPwestAF analyses

---
title: "West Africa cuckoo stopover rainfall patterns"
author: "Mark Miller"
date: "31 July 2018"
output: html_document
---

```{r cuckoos, echo=FALSE, message=F, warning=F, fig.height=9}

#### work around for different laptop/desktop directories ####
#### !!!! RUN THIS !!!! ####
 if(Sys.info()['nodename']=="D9L5812"){
   setwd("C:/cuckoo_tracking")}else{
    setwd("N:/cuckoo_tracking")}

library(ggplot2)
library(dplyr)
#summary data
dat<- read.csv('C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig_summary_dead.csv', h=T)

# set NA or manually correct cuckoo depature dates that
# are incorrect, from manual examination

dat[dat$ptt==115586 & dat$year==2014,]$DEPwestAF<-106
dat[dat$ptt==134955 & dat$year==2015,]$DEPwestAF<-NA
dat[dat$ptt==134956 & dat$year==2015,]$DEPwestAF<-NA

# and dead ones
dat[dat$ptt==128296 & dat$year==2014,]$DEPwestAF<-NA

env<-read.csv('C:/cuckoo_tracking/data/spring_rainfall_NDVI_GRIMMS_TAMSAT_emodis_by_stopover_detailcoords_2018_dead.csv', h=T)

names(env)[names(env)=='value']<-'precip'
names(env)[names(env)=='value2']<-'ndvi'
env$cuck_pres<-as.numeric(env$cuck_pres) # abs=1, pres=2

dat<-left_join(dat, env %>% group_by(ptt, year) %>% arrange(SO_startDOY, variable, .by_group=T) %>%
                 summarise(dep_country=last(country)),
               by=c('ptt', 'year'))

# can test individual countries
#dat<-dat[dat$dep_country=="Cote d'Ivoire",]
#dat<-dat[dat$dep_country=="Ghana",]

# summarise data for each stopover so that value per date is the mean
# of the detail coords values

dat2<- env[,c(2,3,4,5,10:22)] %>% group_by(year, ptt, SO_startDOY, variable) %>%
  summarise_all(mean,na.rm=T)

temp<- env %>% group_by(year, ptt, SO_startDOY, variable) %>% summarise(country=first(country))

dat2$country<-temp$country

dat2[which(dat2$precip>500),]$precip<-0 # remove erroneous vals
dat2[which(is.na(dat2$precip)),]$precip<-0 # fix na

dat2<-dat2 %>% group_by(year, ptt, SO_startDOY) %>% dplyr::mutate(cumrf=cumsum(precip))

# for tamsat cumrf
dat2<-dplyr::arrange(dat2, ptt, year, SO_startDOY, variable)
dat2$tamRAIN[which(!(1:21375 %in% seq(5,21375, 5)))]<-0
dat2<-dat2 %>% group_by(year, ptt, SO_startDOY) %>% dplyr::mutate(tamcumrf=cumsum(tamRAIN))

### visualisastion

for(i in 1:nrow(dat))
{
  dtemp<-dat[i,]

  if(is.na(dtemp$DEPwestAF)){next}

t2<-dat2[dat2$ptt==dtemp$ptt & dat2$year==dtemp$year,]

temp=ggplot(data=t2)+
geom_rect(aes(xmin=SO_startDOY, xmax=SO_endDOY, ymin=0, ymax=500), fill='grey')+

  geom_line( aes(x=variable, y=emodisNDVI*500), colour='dark green')+

    geom_line( aes(x=variable, y=emodisANOM*1000+200), colour='purple', linetype=2)+
  geom_line( aes(x=variable, y=tamRAINanom+200), colour='dark blue', linetype=2)+
  geom_hline(yintercept=200)+
    geom_line(aes(x=variable, y=cumrf), col='blue', size=1)+
  geom_line(aes(x=variable, y=tamcumrf), col='cyan', size=1)+
  facet_wrap(~SO_startDOY+country)+theme_classic()+
  ylab('cumulative rainfall since 01 Jan (mm)')+ xlab('Day of year')+
          ggtitle(paste("PTT:", dtemp$ptt, "Year:",dtemp$year, "Dep WA:", dtemp$DEPwestAF))+ylim(c(0,500))

# for lite version without:
# geom_line( aes(x=variable, y=aquaANOM*1000+200), colour='pink', linetype=2)+
#  geom_line( aes(x=variable, y=aquaNDVI*500), colour='green')+
#
#

print(temp)

}

```