### 2018-08-30 paper_wireframe_knit.Rmd
Removal of final section testing env variables as need to re-think approach.

---
title: "Paper Wireframe"
author: "Mark Miller"
date: "10 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, warning=F, message=F}
library('lme4')
library('car')
library('ggplot2')
library('dplyr')
library('maps')
library('stargazer')
library('splines')
library('sf')
library('leaflet')
library('gridExtra')
library('MuMIn')

if(Sys.info()['nodename']=="D9L5812"){
  setwd("C:/cuckoo_tracking")}else{
    setwd("N:/cuckoo_tracking")}

#summary data
dat<- read.csv('C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig_summary_dead.csv', h=T)

#stopover coord data (median points)
sovers<-read.csv('C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig_BOTH.csv', h=T)

## landcover and river data
lc<-read.csv("C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig_detailcoords_landcov_ext_rivers.csv", h=T)

#extra data for filling migratory route

extraz<- read.csv('C:/cuckoo_tracking/data/stopover_bestofday_1daymin_recalc_spring_mig_summary_extras.csv', h=T)

#### Data cleaning

# set NA or manually correct cuckoo depature dates that
# are incorrect, from manual examination

dat[dat$ptt==115586 & dat$year==2014,]$DEPwestAF<-106
dat[dat$ptt==134955 & dat$year==2015,]$DEPwestAF<-NA
dat[dat$ptt==134956 & dat$year==2015,]$DEPwestAF<-NA

# and dead ones
dat[dat$ptt==128296 & dat$year==2014,]$DEPwestAF<-NA

## fill NA in mig route
# fills some
dat<-left_join(dat, select(extraz, ptt, year, autumn_mig), by=c('ptt', 'year'))

# fill 2018 birds
dat[c(47, 44, 42),]$autumn_mig.y='SW'
dat[39,]$autumn_mig.y='SE'
# and dead birds
dat[c(2,20,45),]$autumn_mig.y='SW'
dat[c(1,3,23,25,31),]$autumn_mig.y='SE'

names(dat)[names(dat)=='autumn_mig.y']<-'autumn_mig'
dat$autumn_mig.x<-NULL

### Add departure info from WA e.g. country, lat and long, length of final stopover and 2nd before last

notWA<-c('Italy', 'France', 'Spain', 'Algeria',
        'United Kingdom', 'Morocco', 'Portugal')

soversWA<-sovers[-which(sovers$country %in% notWA),]

dat<-left_join(dat, soversWA %>% group_by(ptt, year) %>% summarise( dep_country=last(country), dep_long=last(SO_median_long), dep_lat=last(SO_median_lat), st_finalSO=last(SO_startDOY), dur_finalSO_n2=(nth(SO_endDOY, -2)-nth(SO_startDOY, -2))), by=c('ptt', 'year'))

#calc final stopover length (cos we corrected some values in the summary table)

dat$dur_finalSO=dat$DEPwestAF-dat$st_finalSO

#landcover metrics

lc1<-lc %>% group_by(ptt, year, SO_startDOY) %>% dplyr::summarise(Tree=mean(Trees, na.rm=T), Grass=mean(Grasslands, na.rm=T), Crops=mean(Croplands, na.rm=T), Shrub=mean(Shrubs, na.rm=T),Bare=mean(Bare.areas, na.rm=T), Urban=mean(Built.up.areas, na.rm=T),
  Aquatic=mean(Aquatic.veg.reg.flooded, na.rm=T), Water=mean(Open.water, na.rm=T), D_riv=
    mean(d_river), D_riv2=mean(d_river_str2))

# landcover types expressed as a precentage                                                            
lc1[,4:10]<-round(lc1[,4:10]/rowSums(lc1[,4:10], na.rm=T) *100)                             
# join stopover duration to test associations
# lc1 becomes dataset to test stopover length in relation to habitat

lc1<-left_join(lc1, select(soversWA, ptt, year, SO_startDOY, SO_days), by=c('ptt', 'year', 'SO_startDOY'))

## join landcover of final stopover to dat
dat<-left_join(dat, (lc1 %>% group_by(ptt, year) %>% dplyr::summarise_all(last)), by=c('ptt', 'year'))

#### AND THE ENV DATA #####

env<-read.csv('C:/cuckoo_tracking/data/spring_rainfall_NDVI_GRIMMS_TAMSAT_emodis_by_stopover_detailcoords_2018_dead.csv', h=T)

names(env)[names(env)=='value']<-'precip'
names(env)[names(env)=='value2']<-'ndvi'
env$cuck_pres<-as.numeric(env$cuck_pres) # abs=1, pres=2

# summarise data for each stopover so that value per date is the mean
# of the detail coords values

dat2<- env[,c(2,3,4,10:22)] %>% group_by(year, ptt, SO_startDOY, variable) %>%
  summarise_all(mean,na.rm=T)

temp<- env %>% group_by(year, ptt, SO_startDOY, variable) %>% summarise(country=first(country))

dat2$country<-temp$country

dat2[which(dat2$precip>500),]$precip<-0 # remove erroneous vals
dat2[which(is.na(dat2$precip)),]$precip<-0 # fix na

dat2<-dat2 %>% group_by(year, ptt, SO_startDOY) %>% dplyr::mutate(cumrf=cumsum(precip))

# for tamsat cumrf
dat2<-dplyr::arrange(dat2, ptt, year, SO_startDOY, variable)
dat2$tamRAIN[which(!(1:21375 %in% seq(5,21375, 5)))]<-0
dat2<-dat2 %>% group_by(year, ptt, SO_startDOY) %>% dplyr::mutate(tamcumrf=cumsum(tamRAIN))

# select cond during SO..and befoe??
dat2.1<-subset(dat2, cuck_pres==2)

dat2.2<-dat2.1 %>% group_by(year, ptt, SO_startDOY) %>%
  summarise_if(is.numeric, mean, na.rm=T)

# read in model-ready dataset

dmod<-read.csv('C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig_summary_dead_attrib_modelready.csv', h=T)

dmod$tWA=dmod$DEPwestAF-dmod$departureGarr

dmod<-dmod[,c(1,2,6,22,27,28,29,37,39,52,57,63,74,85,87,92, 115)]

dmod<-na.omit(dmod)

```

## Aims

Focus on cuckoo spring departure from West Africa. Address several questions:

Q1) Is cuckoo arrival in the UK constrained by their departure from West Africa ?

Q2) To what extent is cuckoo departure from West Africa controlled by endogenous or environmental factors ?

Q3) What are the mechanisms that underpin potential controlling factors ?

## Methods

## Results

### Q1)

Arrival at UK breeding grounds is less correlated with earlier stages of spring migration, but appears to be still be influenced by movements in West Africa.

```{r, echo=F, fig.cap='Correlation between arrival at UK breeding grounds and stages of spring migration: departure from furthest southerly wintering latitude (pink); Central Africa (cyan); West Africa (blue); North Africa (green); Europe (red); and arrival on UK mainland (black).', warning=F, message=F}

ggplot(data=dat, aes(y=arrive_breeding))+
  geom_point(aes(x=arrive_uk), col=1, shape=1)+
  geom_point(aes(x=DEPeurope), col=2, shape=1)+
  geom_point(aes(x=DEPnorthAF), col=3, shape=1)+
  geom_point(aes(x=DEPwestAF), col=4, shape=1)+
  geom_point(aes(x=DEPcentralAF), col=5, shape=1)+
  geom_point(aes(x=depart_winterSO), col=6, shape=1)+
  xlim(c(-71, 142))+ylab('Arrival at breeding grounds DOY')+
  xlab('DOY')+theme_bw()

```

Stats confirm that date of departure from West Africa significantly impacts subsequent date of arrival at Uk breeding grounds, but prior stages of spring migration do not.

```{r, echo=F}
d1<-dat
  names(d1)[10]<-'arrive.breeding.uk'
  names(d1)[4]<-'depart.wintering'
  names(d1)[5]<-'depart.central.africa'
  names(d1)[6]<-'depart.west.africa'

m1<-lmer(arrive.breeding.uk~depart.wintering+depart.central.africa+depart.west.africa+(1|year)+(1|ptt), data=d1)

Anova(m1, test.statistic='F')

```

For every extra day spent in West Africa cuckoos arrive 0.78 days 
later back at UK breeding grounds.

```{r, warning=F, message=F, echo=F}
m4<-lmer(arrive_breeding~DEPwestAF+
           (1|ptt)+(1|year), data=dat)

newdat<-data.frame(DEPwestAF=70:130, arrive_breeding=1)
newdat$p1<-predict(m4, newdata=newdat, re.form=~0)

predmat<-model.matrix(arrive_breeding~DEPwestAF, data=newdat)
vcv<-vcov(m4)
semod<-sqrt(diag(predmat%*%vcv%*%t(predmat)))
newdat$lc<-newdat$p1-semod*1.96
newdat$uc<-newdat$p1+semod*1.96

qplot(data=dat, x=DEPwestAF, y=arrive_breeding)+
  geom_line(data=newdat, aes(x=DEPwestAF, y=p1), colour='red')+
  geom_line(data=newdat, aes(x=DEPwestAF, y=lc), colour='red', linetype='dashed')+
  geom_line(data=newdat, aes(x=DEPwestAF, y=uc), colour='red', linetype='dashed')+
  xlab("Departure date West Africa (DOY)")+
  ylab("Arrival at breeding grounds (DOY)")+
  theme_bw()
```

Cuckoos that departed West Africa early had longer migrations back to the breeding grounds but those that departed after median departure date (day 102) could not compensate for delayed departure with faster migration.

```{r, warning=F, message=F, echo=F}
d1$migration.length<-d1$arrive.breeding.uk-d1$depart.west.africa

m5<-lmer(migration.length~depart.west.africa+
           (1|ptt)+(1|year), data=d1)

# try segmented model using median departure date as break point
# doent work with mixed model

m5a<-lm(migration.length~depart.west.africa*(depart.west.africa<102)+depart.west.africa*(depart.west.africa>102), data=d1)

m5b<-lmer(migration.length~bs(depart.west.africa, deg=1, knots=102)+
           (1|ptt)+(1|year), data=d1)

newdat<-data.frame(depart.west.africa=70:130, migration.length=1)
newdat$p1<-predict(m5b, newdata=newdat, re.form=~0)

qplot(data=d1, x=depart.west.africa, y=migration.length)+
  xlab("Departure date West Africa (DOY)")+
  ylab("Migration length (days)")+
  theme_bw()+geom_line(data=newdat, aes(x=depart.west.africa, y=p1), colour='red')+geom_label(data=data.frame(), aes(label = 'Anova F=3.54, p=0.046', x = -Inf, y = Inf), hjust = 0, vjust = 1)

#Anova(m5, test.statistic = 'F')
```

## Focus on West Africa

Explore the dataset. Each point gives the final stopover in West Africa before migration. Colours denote different years.

```{r, warning=F, message=F, echo=F}

dat_sp<-st_as_sf(na.omit(dat[,c(1,2,6,23,24)]), coords = c("dep_long","dep_lat"), crs="+proj=longlat +datum=WGS84")

dat_sp$year<-factor(dat_sp$year)

pal <- colorFactor(c("navy", "blue", 'cyan', 'green', 'yellow', 'orange', 'red'), domain = c("2012", "2013", '2014', '2015', '2016', '2017', '2018'))

m <- leaflet(dat_sp) %>%
  addTiles() %>% 
  addCircleMarkers( popup= paste("ptt:", dat_sp$ptt, "<br>", "year:", dat_sp$year,"<br>", "Departure date:", dat_sp$DEPwestAF), color=~pal(year), stroke=F, radius=6, fillOpacity = 0.5)

m

```

### Q2)

## Endogenous factors that influence departure date from West Africa.

No convincing evidence that different breeding populations depart West Africa at different times. *slight significance comes from very early New Forest population*

```{r, warning=F, message=F, echo=F}

#reorder breeding location in latitude order
# relevel to central well sampled year
dat$breeding_loc<-factor(dat$breeding, levels=c('Scotland', 'Lake District', 'North York Moors', 'Sherwood Forest',  'Wales', 'Thetford Forest', 'Norfolk Broads','Ashdown Forest', 'New Forest', 'Dartmoor'))

m1<-lmer(DEPwestAF~breeding_loc+(1|year)+(1|ptt), data=dat)

#Anova(m1, test.statistic = 'F')

ggplot(data=dat, aes(x=breeding_loc, y=DEPwestAF))+geom_jitter(width=0.08, shape=1)+theme_bw()+theme(axis.text.x=element_text(angle=90))+xlab("Breeding location")+ylab("Departure date West Africa (DOY)")+geom_label(data=data.frame(), aes(label = 'Anova F=4.55, p=0.023', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

No evidence that different autumn migration routes impact timing of spring departure from West Africa.

```{r, warning=F, message=F, echo=F}
# and for migratory route
m1<-lmer(DEPwestAF~autumn_mig+(1|year)+(1|ptt), data=dat)

#Anova(m1, test.statistic = 'F')

qplot(data=dat, x=autumn_mig, y=DEPwestAF, geom='boxplot')+xlab("Autumn migratory route")+ylab("Departure date West Africa (DOY)")+geom_label(data=data.frame(), aes(label = 'Anova F=0.81, p=0.373', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

For birds that were tracked through multiple spring migrations there are inter-bird differences in West African departure date and differing levels of inter-annual variation within birds.

```{r, warning=F, message=F, echo=F}
multi_yr<-names(which(table(dat$ptt)>1))

multi_yr<-multi_yr[multi_yr!=128300]

m1<-lmer(DEPwestAF~factor(ptt)+(1|year), data=dat[dat$ptt %in% multi_yr ,])

#Anova(m1, test.statistic = 'F')

ggplot(data=dat[dat$ptt %in% multi_yr ,], aes(x=factor(ptt), y=DEPwestAF ))+xlab("Bird ID")+ylab("Departure date West Africa (DOY)")+geom_jitter(width=0.08,shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=4.13, p=0.017', x = -Inf, y = Inf), hjust = 0, vjust = 1)

```

Differing levels of intra-bird variation were also evident in final stopover location before departure from West Africa

```{r, warning=T, message=T, echo=F}

multi_yr_SO<-sovers[sovers$ptt %in% multi_yr,]

notWA<-c('Italy', 'France', 'Spain', 'Algeria',
        'United Kingdom', 'Morocco')

multi_yr_SO<-multi_yr_SO[-which(multi_yr_SO$country %in% notWA),]

lastWASO<-multi_yr_SO %>% group_by(ptt, year) %>% summarise(long=last(SO_median_long), lat=last(SO_median_lat), country=last(country))

map = map_data("world")

ggplot() + geom_polygon(data = map,
               aes(x=long, y = lat, group = group),fill = NA, colour="darkgray", size=0.5)+
  geom_point(data=lastWASO, aes(x=long, y=lat, colour=factor(ptt)), shape=1, size=2)+
  geom_line(data=lastWASO, aes(x=long, y=lat, colour=factor(ptt)), alpha=0.2)+coord_cartesian()+
   coord_map(xlim = c(-16, 5),ylim = c(14, 3))+xlab('Longitude')+ylab('Latitude')+theme_bw()+ labs(colour = "Bird ID")

```

## Environmental factors that influence departure date from West Africa.

First, there appears no overall difference between years in departure date from West Africa (median: 102 DOY). *This might support internal regulation for timing of migration out of West Africa by birds e.g. photoperiod (endogenous) but could be an artefact of sample size - more tracking would reveal statistically significant differences between years?*

```{r, warning=F, message=F, echo=F}
dat$year<-as.character(dat$year)

m1<-lmer(DEPwestAF~year+(1|ptt), data=dat)

#Anova(m1, test.statistic = 'F')

ggplot(data=dat, aes(x=year, y=DEPwestAF ))+xlab("Year")+ylab("Departure date West Africa (DOY)")+geom_jitter(width=0.08,shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=1.27, p=0.315', x = -Inf, y = Inf), hjust = 0, vjust = 1)

stargazer(m1, type='text',
covariate.labels=c("Constant: 2012","2013", "2014","2015", "2016", "2017", "2018"),
dep.var.labels ="West Africa depature date",
ci=TRUE, intercept.bottom = FALSE, omit.table.layout = "sn", 
digits=1, model.numbers=FALSE,  report='vcs')

```

Country of departure seems to have little overall affect on date of departure but note that birds who leave early (< day 100) all leave from Ivory Coast and (to a lesser extent) Ghana.

```{r, warning=F, message=F, echo=F}

#Anova(lmer(DEPwestAF~dep_country+ (1|ptt)+(1|year), data=dat), test.statistic = 'F')

dtemp<-na.omit(dat[,c(6,22:24)])

dtemp$dep_country<-factor(dtemp$dep_country, levels=c('Sierra Leone', 'Guinea', "Cote d'Ivoire", 'Ghana', 'Burkina Faso', 'Nigeria'))
m1<-lmer(DEPwestAF~breeding_loc+(1|year)+(1|ptt), data=dat)

#Anova(m1, test.statistic = 'F')

ggplot(data=dtemp, aes(x=dep_country, y=DEPwestAF ))+xlab("Departure country")+ylab("Departure date West Africa (DOY)")+geom_jitter(width=0.1,shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=0.552, p=0.735', x = -Inf, y = Inf), hjust = 0, vjust = 1)+theme(axis.text.x=element_text(angle=90))
```

Taking a step back, it is important to if there is any environmental
control over how birds arrive into West Africa.

Drought-breaking rains do not limit departure date specifically, however they do limit arrival into West Africa from Central Africa. Birds do not move into West Africa until rains have arrived. Blue blur is rainfall for each stopover, black line is first arrival of cuckoo and red lines are subsequent stopovers.

```{r, warning=F, message=F, echo=F}

temp<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

temp$country<-factor(temp$country, levels=c('Nigeria', 'Togo', 'Ghana', "Cote d'Ivoire", 'Guinea', 'Sierra Leone'))

ggplot()+geom_line(data=temp[temp$cumrf>0,],aes(x=variable, y=cumrf, group=paste(ptt, year,SO_startDOY)), size=2, alpha=0.08, colour='blue')+theme_bw()+geom_vline(data=(temp %>% group_by(ptt, year, SO_startDOY) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='red', alpha=0.5)+
  geom_vline(data=(temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+facet_wrap(~country)+xlab("Day of year")+
  ylab("Cumulative rainfall (mm)")

```

Shown with mean rainfall (blue line) and NDVI (green line) per country and mean arrival date for first stopovers in West Africa (red line), we can see rains arrive at generally the same time in Cote d'Ivoire and Ghana and so do cuckoos

```{r, warning=F, message=F, echo=F, fig.height=12, fig.width=6}

temp<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

temp$country<-factor(temp$country, levels=c('Nigeria', 'Togo', 'Ghana', "Cote d'Ivoire", 'Guinea', 'Sierra Leone'))

temp2<-temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first) %>% group_by(country) %>% summarise(MSO_startDOY=mean(SO_startDOY), Mrf=mean(cumrf), Mnd=mean(aquaNDVI))

temp3<-temp %>% group_by(country, variable) %>% summarise(Mcumrf=mean(cumrf), rfmin=mean(cumrf)-sd(cumrf), rfmax=mean(cumrf)+sd(cumrf),                                                          MaquaNDVI=mean(aquaNDVI, na.rm=T), ndmin=(mean(aquaNDVI, na.rm=T)-sd(aquaNDVI, na.rm=T)), ndmax=(mean(aquaNDVI, na.rm=T)+sd(aquaNDVI, na.rm=T)))

p1<-ggplot()+
geom_line(data=temp3,aes(x=variable, y=Mcumrf), size=1, colour='blue')+
geom_line(data=temp3,aes(x=variable, y=rfmin), colour='grey', linetype=2)+
geom_line(data=temp3,aes(x=variable, y=rfmax), colour='grey', linetype=2)+  
geom_vline(data=(temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+
geom_vline(data=temp2, aes(xintercept = MSO_startDOY), colour='red', size=1)+
   facet_grid(country~.)+xlab("Day of year")+
  ylab("Cumulative rainfall (mm)")+theme_bw()

p2<-ggplot()+geom_line(data=temp3,aes(x=variable, y=MaquaNDVI), size=1, colour='dark green')+
geom_line(data=temp3,aes(x=variable, y=ndmin), colour='orange', linetype=2)+
geom_line(data=temp3,aes(x=variable, y=ndmax), colour='orange', linetype=2)+
geom_vline(data=(temp[temp$cuck_pres==2,] %>% group_by(ptt, year) %>% summarise_all(first)), aes(xintercept = SO_startDOY), colour='black', alpha=0.5)+
geom_vline(data=temp2, aes(xintercept = MSO_startDOY), colour='red', size=1)+
   facet_grid(country~.)+xlab("Day of year")+
  ylab("NDVI")+theme_bw()

grid.arrange(p1, p2, ncol=2)

```

Despite rain generally arriving in West Africa at the same time, there
is variance of ~ 60 days in the times in which cuckoos arrive into
West Africa (with the bulk of birds arriving between day 55 and day 95)

There is however, little relationship between arrival date into West Africa
and departure date from West Africa

```{r, warning=F, message=F, echo=F}

m1<-lmer(DEPwestAF~departureGarr+(1|ptt)+(1|year), data=dmod[dmod$DEPwestAF%in%85:120,])

ggplot(data=dmod, aes(x=departureGarr, y=DEPwestAF ))+xlab("Arrival into West Africa (DOY)")+ylab("Departure date West Africa (DOY)")+geom_point(shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=1.83, p=0.185 *3 outliers removed*', x = -Inf, y = Inf), hjust = 0, vjust = 1)

#Anova(m1, test.statistic = 'F')

```

Note: if the above model includes the 3 outliers then a there is a significant relationship... * need to decide which is more biologically plausible*

```{r, warning=F, message=F, echo=F}
m1<-lmer(DEPwestAF~departureGarr+(1|ptt)+(1|year), data=dmod)
#Anova(m1, test.statistic = 'F')

newdat<-data.frame(DEPwestAF=1, departureGarr=25:100)
newdat$p1<-predict(m1, newdata=newdat, re.form=~0)

predmat<-model.matrix(DEPwestAF~departureGarr, data=newdat)
vcv<-vcov(m1)
semod<-sqrt(diag(predmat%*%vcv%*%t(predmat)))
newdat$lc<-newdat$p1-semod*1.96
newdat$uc<-newdat$p1+semod*1.96

qplot(data=dmod, y=DEPwestAF, x=departureGarr)+
  geom_line(data=newdat, aes(x=departureGarr, y=p1), colour='red')+
  geom_line(data=newdat, aes(x=departureGarr, y=lc), colour='red', linetype='dashed')+
  geom_line(data=newdat, aes(x=departureGarr, y=uc), colour='red', linetype='dashed')+
  ylab("Days spent in West Africa")+
  xlab("Arrival into West Africa (DOY)")+
  theme_bw()+geom_point(shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=10.47, p=0.0027', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

Because almost all birds depart West Africa between day 90 and day 110,
later arriving birds spend significantly shorter time in West Africa

```{r, warning=F, message=F, echo=F}
m1<-lmer(tWA~departureGarr+(1|year)+(1|ptt), data=dmod)

#Anova(m1, test.statistic = 'F')

newdat<-data.frame(tWA=1, departureGarr=25:100)
newdat$p1<-predict(m1, newdata=newdat, re.form=~0)

predmat<-model.matrix(tWA~departureGarr, data=newdat)
vcv<-vcov(m1)
semod<-sqrt(diag(predmat%*%vcv%*%t(predmat)))
newdat$lc<-newdat$p1-semod*1.96
newdat$uc<-newdat$p1+semod*1.96

qplot(data=dmod, y=tWA, x=departureGarr)+
  geom_line(data=newdat, aes(x=departureGarr, y=p1), colour='red')+
  geom_line(data=newdat, aes(x=departureGarr, y=lc), colour='red', linetype='dashed')+
  geom_line(data=newdat, aes(x=departureGarr, y=uc), colour='red', linetype='dashed')+
  ylab("Days spent in West Africa")+
  xlab("Arrival into West Africa (DOY)")+
  theme_bw()+geom_point(shape=1)+theme_bw()+geom_label(data=data.frame(), aes(label = 'Anova F=63.35, p=3.19e-09', x = -Inf, y = Inf), hjust = 0, vjust = 1)
```

When graphed with average rainfall and ndvi curved (across all WA stopovers) we see that early arriving birds into West Africa spend a long time there when greening is yet to/just starting whereas late arriving birds spend a short time at near-peak greenness. Interestingly,
the mean arrival date into west africa (day 71; black line) corresponds with earlier departures, which could be facilitated by concurrent peak greening and rainfall rates.

```{r, warning=F, message=F, echo=F}
temp<- dat2[- which(dat2$country %in% c('Angola', 'Central African Republic','Congo', 'Democratic Republic of the Congo', 'Gabon',  'Cameroon', 'Burkina Faso', 'Liberia')),]

temp3<-temp %>% group_by(variable) %>%
  summarise(Mcumrf=mean(tamcumrf), rfmin=mean(tamcumrf)-sd(tamcumrf), rfmax=mean(tamcumrf)+sd(tamcumrf),
   MaquaNDVI=mean(emodisNDVI, na.rm=T), ndmin=(mean(emodisNDVI, na.rm=T)-sd(emodisNDVI, na.rm=T)),
   ndmax=(mean(emodisNDVI, na.rm=T)+sd(emodisNDVI, na.rm=T)))

#lets make variable!

ggplot(data=dmod, aes(x=departureGarr, y=DEPwestAF))+
  geom_line(data=temp3,aes(x=variable, y=Mcumrf/2), size=1, colour='blue')+
  geom_line(data=temp3,aes(x=variable, y=rfmin/2), colour='grey', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=rfmax/2), colour='grey', linetype=2)+ 
  geom_line(data=temp3,aes(x=variable, y=MaquaNDVI*100), size=1, colour='dark green')+
  geom_line(data=temp3,aes(x=variable, y=ndmin*100), colour='orange', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=ndmax*100), colour='orange', linetype=2)+
  geom_point(aes(colour=tWA))+scale_y_continuous(breaks=seq(0, 130, 10), limits=c(0,130))+
  scale_x_continuous(breaks=seq(0, 130, 10))+
  scale_colour_gradientn(colours=topo.colors(8))+
  theme_bw()+labs(y="Departure from West Africa (DOY)", x="Arrival into West Africa (DOY)", colour='Days in WA')+geom_vline(aes(xintercept=71), colour='black')
```

Visualised another way with time coloured gradient.

```{r, warning=F, message=F, echo=F} 
df <- expand.grid(departureGarr=0:130,
                  DEPwestAF=75:130)

df$filly=df$DEPwestAF-df$departureGarr

df[df$filly<0 | df$filly>80,]$filly<-NA

ggplot() + 
  geom_tile(data=df, aes(x=departureGarr, y=DEPwestAF, fill=filly),alpha = 0.75) +      
  scale_fill_gradient(low='#ffeda0', high='#f03b20', na.value=NA)+
  geom_line(data=temp3,aes(x=variable, y=Mcumrf/2), size=1, colour='blue')+
  geom_line(data=temp3,aes(x=variable, y=rfmin/2), colour='grey', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=rfmax/2), colour='grey', linetype=2)+ 
  geom_line(data=temp3,aes(x=variable, y=MaquaNDVI*100), size=1, colour='dark green')+
  geom_line(data=temp3,aes(x=variable, y=ndmin*100), colour='orange', linetype=2)+
  geom_line(data=temp3,aes(x=variable, y=ndmax*100), colour='orange', linetype=2)+
  geom_point(data=dmod, aes(x=departureGarr, y=DEPwestAF), shape=1)+
  theme_bw()+scale_y_continuous(breaks=seq(0, 130, 10), limits=c(0,130))+
  scale_x_continuous(breaks=seq(0, 130, 10))+
  labs(y="Departure from West Africa (DOY)", x="Arrival into West Africa (DOY)", fill='Days in WA')+geom_vline(aes(xintercept=71), colour='black')

```