### 2018-09-17 gd_cuck_plots.Rmd

---
title: "Good Cuck movements"
author: "MM"
date: "13/03/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cuckoos, echo=FALSE, warning=FALSE, message=FALSE}
########################

rm(list=ls())
library(maps)
library(sp)
library(maptools)
library(rgdal)
library(ggplot2)
library(ggmap)
library(ggrepel)

# using best of day data

dat<-read.csv("C:/cuckoo_tracking/data/stopover_bestofday_2018_1daymin_recalc_spring_mig.csv", h=T)

dat2<-read.csv("C:/cuckoo_tracking/data/processed_movebank_cuckoos_hybrid_filter_2018_clean_stopovers_recalc.csv", h=T)

UK_travel<-read.csv("C:/cuckoo_tracking/data/complete_cycles_UK_all_timing.csv", h=T)

UK_travel<-UK_travel[UK_travel$deployment_entry==FALSE,]

dat2$key<-paste(dat2$ptt, dat2$year.x)

dat3<-dat2[dat2$key %in% unique(paste(UK_travel$ptt, UK_travel$year)),]

dat3<-dat3[dat3$month<6,]

names(dat3)[names(dat3)=='year.x']<-'year'

m1 <- get_map(location = c(5, 30), zoom=4)

for(i in unique(dat3$ptt))
{
  cols <- c("2012" = "red", "2013" = "blue", "2014" = "darkgreen", "2015" = "orange", "2016"= "cyan", "2017"= "black",
            "2018"= "purple")

  g<-ggmap(m1)+
    geom_path(data=dat3[dat3$ptt==i,], aes(x=long, y=lat,                                      colour=factor(year)))+
    geom_jitter(data=dat3[dat3$ptt==i,], aes(x=long, y=lat,
                colour=factor(year)), shape=1, width = 0.5, height= 0.5)+ggtitle(i)+

  geom_label_repel(data=dat[dat$ptt==i & dat$deployment_entry==FALSE,], aes(x=SO_median_long, y=SO_median_lat,
                colour=factor(year), label=substr(SO_start, 6,10)))+
  geom_label_repel(data=aggregate(year~breeding_entry, data=dat[dat$ptt==i,], FUN=function(x){unique(x)})
                   , aes(x=-18, y=45,
                colour=factor(year), label=substr(breeding_entry, 6,10))) + scale_colour_manual(values = cols)

  print(g)

}

```